### 2013-12-27 draft/references.tex
Add reference.tex. Fill in fields overlapped with Wright 2011 in table.

% List references here
\begin{thebibliography}

\bibitem[Wright et al.(2011)]{Wright2011} Wright, J.~T., et al.\ 2011, \pasp, 123, 412

\end{thebibliography}