### 2014-01-02 draft/references.tex
Minor content revisions, add references.

% List references here
\begin{thebibliography}

\bibitem[Butler et al.(2002)]{Butler2002} Butler, R.~P., et al.\ 2002, 
\apj, 578, 565

\bibitem[Butler et al.(2006)]{Butler2006} Butler, R.~P., et al.\ 2006, 
\apj, 646, 505

\bibitem[Schneider et al.(2011)]{Schneider2011} Schneider, J., Dedieu, C., 
Le Sidaner, P., Savalle, R., \& Zolotukhin, I.\ 2011, \aap, 532, A79

\bibitem[Wright et al.(2011)]{Wright2011} Wright, J.~T., et al.\ 2011, \pasp, 123, 412

\end{thebibliography}