### 2014-02-18 draft/references.tex

% List references here
\begin{thebibliography}

\bibitem[Akeson et al.(2013)]{Akeson2013} Akeson, R.~L., et al.\ 2013, 
\pasp, 125, 989 % NASA Exoplanet Archive

\bibitem[Batalha et al.(2013)]{Batalha2013} Batalha, N.~M., et al.\ 2013, 
\apjs, 204, 24 % example of KOI release

\bibitem[Borucki et al.(2010)]{Borucki2010} Borucki, W.~J., et al.\ 2010, 
Science, 327, 977 % Kepler instroduction and first results

\bibitem[Butler et al.(2002)]{Butler2002} Butler, R.~P., et al.\ 2002, 
\apj, 578, 565 % first peer-reviewed list of planets with robust orbits

\bibitem[Butler et al.(2006)]{Butler2006} Butler, R.~P., et al.\ 2006, 
\apj, 646, 505 % EOD first paper

\bibitem[Schneider et al.(2011)]{Schneider2011} Schneider, J., Dedieu, C., 
Le Sidaner, P., Savalle, R., \& Zolotukhin, I.\ 2011, \aap, 532, A79
% exoplanet.eu

\bibitem[Wright et al.(2011)]{Wright2011} Wright, J.~T., et al.\ 2011,
  \pasp, 123, 412 % EOD 2011 paper

\end{thebibliography}