### 2014-03-06 draft/references.tex

% List references here
\begin{thebibliography}

\bibitem[Akeson et al.(2013)]{Akeson2013} Akeson, R.~L., et al.\ 2013, 
\pasp, 125, 989 % NASA Exoplanet Archive

\bibitem[Batalha et al.(2013)]{Batalha2013} Batalha, N.~M., et al.\ 2013, 
\apjs, 204, 24 % example of KOI release

\bibitem[Borucki et al.(2010)]{Borucki2010} Borucki, W.~J., et al.\ 2010, 
Science, 327, 977 % Kepler instroduction and first results

\bibitem[Butler et al.(2002)]{Butler2002} Butler, R.~P., et al.\ 2002, 
\apj, 578, 565 % first peer-reviewed list of planets with robust orbits

\bibitem[Butler et al.(2006)]{Butler2006} Butler, R.~P., et al.\ 2006, 
\apj, 646, 505 % EOD first paper

\bibitem[Cosentino et al.(2012)]{Cosentino2012} Cosentino, R., et al.\ 
2012, \procspie, 8446, 1V % HARPS-North

\bibitem[Dawson \& Murray-Clay(2013)]{Dawson2013} Dawson, R.~I., \&
  Murray-Clay, R.~A.\ 2013, \apjl, 767, L24  % use of EOD

\bibitem[Fabrycky \& Winn(2009)]{Fabrycky2009} Fabrycky, D.~C., \&
  Winn, J.~N.\ 2009, \apj, 696, 1230 % spin-orbit misalignment lambda definition

\bibitem[Green et al.(2012)]{Green2012} Green, J., et al.\ 2012, 
arXiv:1208.4012   % WFIRST

\bibitem[Howard(2013)]{Howard2013} Howard, A.~W.\ 2013, Science, 340,
  572 % use of EOD

\bibitem[Kane(2014)]{Kane2014} Kane, S.~R.\ 2014, \apj, 782, 111 % use
                                % of EOD

\bibitem[Kipping(2013)]{Kipping2013} Kipping, D.~M.\ 2013, \mnras,
  434, L51 % use of EOD

\bibitem[Lissauer et al.(2011)]{Lissauer2011} Lissauer, J.~J., et al.\ 
2011, \apjs, 197, 8 % Mass-Radius for 1 Re < R < 6 Re

\bibitem[Macintosh et al.(2014)]{Macintosh2014} Macintosh, B., Gemini 
Planet Imager instrument team, Planet Imager Exoplanet Survey, G., 
\& Observatory, G.\ 2014, American Astronomical Society Meeting
Abstracts, 223, \#229.02  % GPI

\bibitem[Marcy et al.(2014)]{Marcy2014} Marcy, G.~W., et al.\ 2014, \apjs, 
210, 20 % 42 RV confirmed small Kepler planets

\bibitem[Mordasini et al.(2012)]{Mordasini2012} Mordasini, C., Alibert, Y., 
Georgy, C., Dittkrist, K.-M., Klahr, H., \& Henning, T.\ 2012, \aap, 547, A112
% Mass-Radius for 6 Re < R < 11.684 Re

\bibitem[Ricker et al.(2010)]{Ricker2010} Ricker, G.~R., et al.\ 2010, 
Bulletin of the American Astronomical Society, 42, \#450.06 % TESS

\bibitem[Rowe et al.(2014)]{Rowe2014} Rowe, J.~F., et al.\ 2014, 
arXiv:1402.6534 % 715 new Kepler planets (multis)

\bibitem[Schneider et al.(2011)]{Schneider2011} Schneider, J., Dedieu, C., 
Le Sidaner, P., Savalle, R., \& Zolotukhin, I.\ 2011, \aap, 532, A79
% exoplanet.eu

\bibitem[Skrutskie et al.(2006)]{Skrutskie2006} Skrutskie, M.~F., et al.\ 
2006, \aj, 131, 1163 % 2MASS

\bibitem[Sotin et al.(2007)]{Sotin2007} Sotin, C., Grasset, O., 
\& Mocquet, A.\ 2007, \icarus, 191, 337 % Mass-Radius for M<1Me, R<1Re

\bibitem[van Leeuwen(2009)]{van Leeuwen2009} van Leeuwen, F.\ 2009, \aap, 
497, 209 % Hipparcos catalog

\bibitem[Weiss \& Marcy(2014)]{Weiss2014} Weiss, L.~M., \& Marcy,
  G.~W.\ 2014, \apjl, 783, L6  % Mass-Radius for R < 4Re to be implemented

\bibitem[Wright et al.(2011)]{Wright2011} Wright, J.~T., et al.\ 2011,
  \pasp, 123, 412 % EOD 2011 paper

\bibitem[Wright \& Gaudi(2013)]{Wright2013} Wright, J.~T., \& Gaudi, B.~S.\ 2013,
  Planets, Stars and Stellar Systems.~Volume 3: Solar and Stellar
  Planetary Systems, 489 % Review article with definition of orbital elements

\bibitem[Wright et al.(2014)]{Wright2014} Wright, J., et al.\ 2014, 
American Astronomical Society Meeting Abstracts, 223, \#148.31

\end{thebibliography}