### 2014-09-29 arxiv_newsletter/coolnews_han2014.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                                                       %%%
%%%            LaTeX MACRO FOR COOLNEWS (Journal Abstracts)               %%%
%%%                                                                       %%%
%%%    Simply fill in the information between the brackets { } below      %%%
%%%    and e-mail to coolnews@jila.colorado.edu.  Try not to use          %%%
%%%    specially-defined Latex commands, but if you must then include     %%% 
%%%    their definitions.                                                 %%%
%%%                                                                       %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% To run Latex on your journal abstract, first remove the % sign in the
%% nine lines below, and in the \end{document} line.

%\documentstyle{article}
%\textwidth 18cm
%\textheight 23cm
%\oddsidemargin -1cm
%\topmargin 0cm
%\parskip 0.15cm
%\parindent 0pt
%\small
%\begin{document}

%%            ------ TITLE ---
%% Write the title of your paper between the brackets. Please
%% capitalize only the first letter of each word.
{\large\bf{The Exoplanet Orbit Database \textsc{II}: Updates to
  exoplanets.org}

%%            ------ AUTHORS -----
%% Here comes the author(s) of the paper, please indicate within $^...$ 
%% the number which corresponds to the institute of each author.
{\bf{Eunkyu Han$^{1,2}$, Sharon X.~Wang$^{1,2,3}$, Jason T.~Wright$^{1,2,3}$, Y.\
  Katherian Feng$^{1,2}$, Ming Zhao$^{1,2}$, Onsi
  Fakhouri$^{4}$, Jacob I.\ Brown$^{1,2}$, Colin Hancock$^{1,2}$}}

%%           ------INSTITUTIONS ---
%% Here write your institute name(s) and address(es), 
%% the number in $^..$ indicates your author number, for example:
$^1$ {Department of Astronomy \& Astrophysics, 525 Davey
  Lab, The Pennsylvania State University, University Park, PA 16802, USA}
$^2$ {Center for Exoplanets and Habitable Worlds, The
  Pennsylvania State University, University Park, PA 16802, USA}
$^3$ {Penn State Astrobiology Research Center, The
  Pennsylvania State University, University Park, PA 16802, USA}
$^4$ {Pivotal Labs, Cloud Foundry, 875 Howard St., Fourth
  Floor, San Francisco, CA 94103}

%%            ------ABSTRACT------
%% Enter the text of your abstract between these brackets:
{The Exoplanet Orbit Database (EOD) compiles orbital, transit, host
star, and other parameters of robustly detected exoplanets reported in
the peer-reviewed literature. The EOD can be navigated through the
Exoplanet Data Explorer (EDE) Plotter and Table, available on the
World Wide Web at exoplanets.org. The EOD contains data for 1492
confirmed exoplanets as of July 2014.  The EOD descends from a table
in Butler et al.\ 2002 and the Catalog of Nearby Exoplanets
Butler et al.\ 2006, and the first complete documentation for the EOD
and the EDE was presented in Wright et al.\ 2011. In this work, we
describe our work since then.  We have expanded the scope of the EOD
to include secondary eclipse parameters, asymmetric uncertainties, and
expanded the EDE to include the sample of over 3000 \textit{Kepler} Objects
of Interest (KOIs), and other real planets without good orbital
parameters (such as many of those detected by microlensing and
imaging). Users can download the latest version of the entire EOD as a
single comma separated value file from the front page of
exoplanets.org.}

%%            -----PUBLICATION STATUS---
%% Here give the  publication status (Accepted by, Submitted to) 
%% and the name of journal, for example:
{ PASP in press, arXiv: }

%%            -----E-MAIL ADDRESS-----
%% Here give the e-mail address for preprint requests, for example:
{{\em For preprints contact}: jtwright@astro.psu.edu}
%%

%%            ----FTP or WWW ADDRESS (optional)----
%% If your preprint is available via ftp or WWW, uncomment
%% the line below by removing the % sign in first column
%% and enter the ftp or WWW address.

%{{\em For preprints via ftp or WWW}: http://www.myaddress } 
%% 
%%             ----CATEGORY------
%% Below give the subject category of the abstract (stellar, solar).
%% If you aren't sure, then give Category: uncertain.  
%% The category will only be used for sorting, and wont be printed     
%% in the newsletter.
{ Category: stellar }
%
%\vspace*{0.3cm}
%\end{document}